---
title: "Best Places for Heart Disease?"
author: "Aaron Niskin, Chris Leonard, Lance Price, Nicole Navarro"
date: "November 30, 2016"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE}
library(tidyverse)
library(xtable)
setwd("~/Documents/courses/newCollege/current/eda/projects/project2/")
source("http://peterhaschke.com/Code/multiplot.R")

merged_best <- read.csv("csv_files/merged_best.csv")
merged_worst <- read.csv("csv_files/merged_worst.csv")
```

##Introduction slide?

Question to answer:

>- Where are the best and worst places in the US to have a heart attack?


##There is Another Side to The Problem

It is estimated that each year in the United States:

>- 326,200 people experience EMS-assessed out-of-hospital cardiac arrests

>- 209,000 people are treated for in-hospital cardiac arrest.

>- Approximately 10.6% of people who experience nontraumatic EMS-treated cardiac arrest survive to discharge

##Time is of The Essence

Brain death and permanent death begin to occur in 4-6 minutes after someone experiences cardiac arrest.

EMS response time is a decisive factor in out-of-hospital cardiac arrests.

>- response time data is not recorded on a state or national level

>- it is left up to individual EMS agencies or counties to make decide what to keep track of



##EMS Response Times in Los Angeles

```{r, warning=FALSE,message=FALSE}
LATimes <- read.csv("csv_files/LA_response_times.csv")
ggplot(LATimes, aes(Minutes)) + geom_histogram(bins=50) +
  scale_x_continuous(breaks = round(seq(min(LATimes$Minutes), max(LATimes$Minutes), by = 5),1))
```

##EMS Response Times in New York

```{r, warning=FALSE,message=FALSE}
NYTimes <- read.csv("csv_files/NY_response_times.csv")
ggplot(NYTimes, aes(Minutes)) + geom_histogram(bins=50) +
  scale_x_continuous(breaks = round(seq(min(LATimes$Minutes), max(LATimes$Minutes), by = 5),1))
```



<!-- Chris's part -->

```{r}
complications <- read.csv("csv_files/main_complications.csv")
complications <- complications[,2:dim(complications)[[2]]]
names(complications)[1] <- "County"
```


## Complications Post Surgery

## Complications Post Surgery
- Servere Blood Clots


## Complications Post Surgery
- Servere Blood Clots
- Blood Stream Infection


## Complications Post Surgery
- Servere Blood Clots
- Blood Stream Infection
- Infection of Catheter


## Complications Post Surgery
- Servere Blood Clots
- Blood Stream Infection
- Infection of Catheter
- Other Serious Complications


```{r, echo=FALSE} 
##For top 20!
clot_desc <- filter(complications, Score.Clot > 12.10)
sepsis_desc <- filter(complications, Score.Sepsis>17.44)
cath_desc <- filter(complications, Score.Cath.Infection >.49)
complications_desc <- filter(complications, Score.Serious.Complication > 1.52)



clot_asc <- filter(complications, Score.Clot < 2.48)
sepsis_asc <- filter(complications, Score.Sepsis < 7.16)
cath_asc <- filter(complications, Score.Cath.Infection <.10)
complications_asc <- filter(complications, Score.Serious.Complication < .64)
```

##

```{r,echo=FALSE}
ggplot(clot_asc, aes(x=reorder(County,Score.Clot),y=Score.Clot, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Lowest Blood Clot Probability") + xlab("Top Counties") + ylab("Patient Probability of Clot")
```

##

```{r,echo=FALSE}
ggplot(clot_desc, aes(x=reorder(County,Score.Clot),y=Score.Clot, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Highest Blood Clot Probability") + xlab("Top Counties") + ylab("Patient Probability of Clot")
```

##

```{r,echo=FALSE}
ggplot(sepsis_asc, aes(x=reorder(County,Score.Sepsis),y=Score.Sepsis, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Lowest Sepsis Probability") + xlab("Top Counties") + ylab("Patient Probability of Bacterial Sepsis")
```

##

```{r,echo=FALSE}
ggplot(sepsis_desc, aes(x=reorder(County,Score.Sepsis),y=Score.Sepsis, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Highest Sepsis Probability") + xlab("Top Counties") + ylab("Patient Probability of Bacterial Sepsis")
```

##

```{r, echo=FALSE}
ggplot(cath_asc, aes(x=reorder(County,Score.Cath.Infection),y=Score.Cath.Infection, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Lowest Catheder Infection Probability") + xlab("Top Counties") + ylab("Patient Probability of Infection By Catheder")
```

##

```{r,echo=FALSE}
ggplot(cath_desc, aes(x=reorder(County,Score.Cath.Infection),y=Score.Cath.Infection, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Highest Catheder Infection Probability") + xlab("Top Counties") + ylab("Patient Probability of Infection By Catheder")
```

##

```{r,echo=FALSE}
ggplot(complications_asc, aes(x=reorder(County,Score.Serious.Complication),y=Score.Serious.Complication, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Lowest Severe Complications in Surgery") + xlab("Top Counties") + ylab("Patient Probability of Complications in Surgery")
```

##

```{r,echo=FALSE}
ggplot(complications_desc, aes(x=reorder(County,Score.Serious.Complication),y=Score.Serious.Complication, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Highest Severe Complications in Surgery") + xlab("Top Counties") + ylab("Patient Probability of Complications in Surgery")
```

## Next Step
- What if we have other data on health?

  * Expanding beyond top 20

```{r, eval=FALSE}
clot_quants <- quantile(complications$Score.Clot, na.rm = TRUE)
best_clots <- complications[!is.na(complications$Score.Clot) & complications$Score.Clot <= clot_quants[[2]], ]
worst_clots <- complications[!is.na(complications$Score.Clot) & complications$Score.Clot >= clot_quants[[4]], ]
rm(clot_quants)

sepsis_quants <- quantile(complications$Score.Sepsis, na.rm = TRUE)
best_seps <- complications[!is.na(complications$Score.Sepsis) & complications$Score.Sepsis <= sepsis_quants[[2]], ]
worst_seps <- complications[!is.na(complications$Score.Sepsis) & complications$Score.Sepsis >= sepsis_quants[[4]], ]
rm(sepsis_quants)
```

<!-- Lance's part -->

##Mortality Rate

```{r echo = FALSE}
H_data1_ami_mort <- read.csv("csv_files/H_data1_ami_mort.csv")

H_data1_ami_readm <- read.csv("csv_files/H_data1_ami_readm.csv")


ggplot(H_data1_ami_mort, aes(Death_Score)) + geom_density(na.rm=TRUE)
```

##Readmission Rate

```{r echo = FALSE echo = FALSE}
ggplot(H_data1_ami_readm, aes(Readmission_Score)) + geom_density(na.rm=TRUE)

```

##Mortality Rate

```{r echo = FALSE}
boxplot(H_data1_ami_mort$Death_Score, main = "Mortality Rate")
```

##Readmission Rate

```{r echo = FALSE}
boxplot(H_data1_ami_readm$Readmission_Score, main = "Readmission Rate")
```

##Mortality VS Readmission (Average Hospital per County)

```{r echo = FALSE}
counties_avg_scores <- read.csv("csv_files/counties_avg_scores.csv")

hospitals_scores <- read.csv("csv_files/hospitals_scores.csv")

ggplot(counties_avg_scores, aes(Average_Mort_Score, Average_Readm_Score)) + geom_point(aes(color=as.factor(Death_Readm_Cat)), na.rm=TRUE) + geom_smooth(method = lm, na.rm = TRUE) 
```

##Mortality VS Readmission per Hospital

```{r echo = FALSE}
ggplot(hospitals_scores, aes(Death_Score, Readmission_Score)) + geom_point(aes(color=as.factor(Death_Readm_Cat)), na.rm=TRUE) + geom_smooth(method = lm, na.rm = TRUE)
```

##Percent of Good Hospitals Vs Total Hospitals Per County

```{r echo = FALSE}
counties_best_mort_avg_filt <- read.csv("csv_files/counties_best_mort_avg_filt.csv")

ggplot(counties_best_mort_avg_filt, aes(percent_good, Count_Total_Hospitals)) + geom_point(aes(color=as.factor(County)))
```

##Percent of Bad Hospitals Vs. Total Hospitals Per County

```{r,  echo = FALSE, warning=FALSE}
counties_worst_mort_avg_filt <- read.csv("csv_files/counties_worst_mort_avg_filt.csv")

ggplot(counties_worst_mort_avg_filt, aes(percent_bad, Count_Total_Hospitals)) + geom_point(aes(color=as.factor(County)))
```

```{r, eval=FALSE}
best_mort <- read.csv("csv_files/counties_best_mort_avg.csv")
worst_mort <- read.csv("csv_files/counties_worst_mort_avg.csv")
```

<!-- Aaron's part -->
##Prevention

>-  Factors
    1.  Nutrition
    2.  Smoking
    
##Nutrition

```{r, warning=FALSE, results='hold', message=FALSE}
smoking <- read.csv("csv_files/smoking.csv", stringsAsFactors = FALSE)
aggriData <- read.csv("csv_files/aggriculture_aggregated.csv", stringsAsFactors = FALSE)
names(aggriData) <- gsub("\\.", ",", gsub("_", "\\ ", names(aggriData)))
smoking <- smoking %>% mutate(County = toupper(County))
aggriData <- aggriData %>% mutate(County = toupper(County))
```

```{r, eval=FALSE, warning=FALSE}
smoking_quants <- quantile(smoking$percent_smoking, na.rm = TRUE)
best_smoking <- smoking[smoking$percent_smoking <= smoking_quants[[2]], ]
worst_smoking <- smoking[smoking$percent_smoking >= smoking_quants[[4]],]
rm(smoking_quants)
merged_best <- inner_join(best_clots, best_smoking, by=c("County", "State"))
merged_best <- inner_join(merged_best, best_mort, by=c("County", "State"))
merged_best <- inner_join(merged_best, best_seps)
merged_worst <- inner_join(worst_clots, worst_smoking, by=c("County", "State"))
merged_worst <- inner_join(merged_worst, worst_mort, by=c("County", "State"))
merged_worst <- inner_join(merged_worst, worst_seps)
write.csv(merged_best, "csv_files/merged_best.csv", row.names = FALSE)
write.csv(merged_worst, "csv_files/merged_worst.csv", row.names = FALSE)
```

```{r, warning=FALSE}
dats <- inner_join(smoking, aggriData, by=c("County", "State"))
dats_10 <- dats[,c(1, 2, 3, 9)]
names(dats_10) <- c("State", "County", "Percent_Smoking", "Low_Access_10_Miles_Share")
```

```{r, eval=FALSE, warning=FALSE, message=FALSE}
smoking_quants <- quantile(dats_10$Percent_Smoking)
worst_smoking <- smoking[smoking$percent_smoking >= smoking_quants[[4]], ]
best_smoking <- smoking[smoking$percent_smoking <= smoking_quants[[2]], ]

food_quants <- quantile(aggriData$`Low access, seniors age 65, at 10 miles, share`)
worst_food <- aggriData[aggriData$`Low access, seniors age 65, at 10 miles, share` >= food_quants[[3]],]
best_food <- aggriData[aggriData$`Low access, seniors age 65, at 10 miles, share` <= food_quants[[2]],]
best_counties_food_and_smoking <- inner_join(best_food, best_smoking, by=c("State", "County")) %>% mutate(County = toupper(County))
worst_counties_food_and_smoking <- inner_join(worst_food, worst_smoking, by=c("State", "County")) %>% mutate(County = toupper(County))
```

```{r, results='show', warning=FALSE, message=FALSE}
merged_best = inner_join(merged_best, aggriData)
merged_worst = inner_join(merged_worst, aggriData)
ggplot(data=dats_10, aes(Low_Access_10_Miles_Share)) + geom_density(fill='tomato2') + labs(title="Density Plot for Low Access to Food", x="Percentage with Low Access") + geom_vline(data= merged_worst, aes(xintercept = `Low access, seniors age 65, at 10 miles, share`)) + geom_vline(data = merged_best, aes(xintercept = `Low access, seniors age 65, at 10 miles, share`, color=County))
```

##Smoking

```{r, results='show', warning=FALSE}
ggplot(data=dats_10, aes(Percent_Smoking)) + geom_density(fill='turquoise3') + labs(x="Percentage of Smokers in Population") + geom_vline(aes(xintercept = merged_worst$percent_smoking)) + geom_vline(data = merged_best, aes(xintercept = percent_smoking, color=County))
```


<!--
##Relationship Between Nutrition And Smoking


```{r, results='show'}
ggplot(data=dats_10, aes(x=Low_Access_10_Miles_Share, y=Percent_Smoking)) + geom_point(alpha=0.8, size=0.2) + geom_smooth() + labs(x="Poor Access to Supermarkets", y="Percentage of Smokers in Population", title="Smoking Percentage vs Access to Nutrition")
```
-->

<!-- ##Obesity -->

```{r, message=FALSE, warning=FALSE}
countyHealth <- read.csv("csv_files/countyHealthRankings.csv")
countyHealth$County <- toupper(countyHealth$County)
merged_best <- inner_join(merged_best, countyHealth)
merged_worst <- inner_join(merged_worst, countyHealth)
# ggplot(data=countyHealth, aes(x=Obesity)) + geom_density(fill='darkolivegreen2') + labs(title="Percentage of Obesity in Population") + geom_vline(aes(xintercept = merged_worst$Obesity)) + geom_vline(data = merged_best, aes(xintercept = Obesity, color=County))
```


##Physical Inactivity VS Obesity
```{r, message=FALSE, warning=FALSE}
# ggplot(data=countyHealth, aes(x=Chlamydia)) + geom_density(fill='darkolivegreen2') + labs(title="Percentage of Chlamydia in Population") + geom_vline(aes(xintercept = merged_worst$Chlamydia)) + geom_vline(data = merged_best, aes(xintercept = Chlamydia, color=County))
# ggplot(data=countyHealth, aes(x= Excessive_Drinking, y=Chlamydia)) + geom_point(fill='darkolivegreen2') + labs(title="Chlamyida VS Excessive Drinking") + geom_vline(aes(xintercept = merged_worst$Excessive_Drinking)) + geom_vline(data = merged_best, aes(xintercept = Excessive_Drinking, color=County))
ggplot(data=countyHealth, aes(x= Obesity, y=Physically_Inactive)) + geom_point(fill='darkolivegreen2', color='lightblue') + labs(title="Physical Inactivity VS Obesity") + geom_point(data = merged_best, aes(x=Obesity, y=Physically_Inactive, color=County)) + geom_point(data=merged_worst, aes(x=Obesity, y=Physically_Inactive, color=County)) 
# ggplot(data=countyHealth, aes(x= Chlamydia, y=Physically_Inactive)) + geom_point(fill='darkolivegreen2', color='lightblue') 
```

##References
