---
title: "Untitled"
author: "Chris Leonard"
date: "December 1, 2016"
output: html_document
---

## Complications Post Surgery

## Complications Post Surgery
- Servere Blood Clots


## Complications Post Surgery
- Servere Blood Clots
- Blood Stream Infection


## Complications Post Surgery
- Servere Blood Clots
- Blood Stream Infection
- Infection of Catheter


## Complications Post Surgery
- Servere Blood Clots
- Blood Stream Infection
- Infection of Catheter
- Other Serious Complications


```{r, echo=FALSE} 
##For top 20!
clot_desc <- dplyr::filter(main_data, Score.Clot > 12.10)
sepsis_desc <- dplyr::filter(main_data, Score.Sepsis>17.44)
cath_desc <- dplyr::filter(main_data, Score.Cath.Infection >.49)
complications_desc <- dplyr::filter(main_data, Score.Serious.Complication > 1.52)



clot_asc <- dplyr::filter(main_data, Score.Clot < 2.48)
sepsis_asc <- dplyr::filter(main_data, Score.Sepsis < 7.16)
cath_asc <- dplyr::filter(main_data, Score.Cath.Infection <.10)
complications_asc <- dplyr::filter(main_data, Score.Serious.Complication < .64)
```

##1

```{r,echo=FALSE}
library(ggplot2)

ggplot(clot_asc, aes(x=reorder(County,Score.Clot),y=Score.Clot, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Lowest Blood Clot Probability") + xlab("Top Counties") + ylab("Patient Probability of Clot")
```

##2

```{r,echo=FALSE}
ggplot(clot_desc, aes(x=reorder(County,Score.Clot),y=Score.Clot, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Highest Blood Clot Probability") + xlab("Top Counties") + ylab("Patient Probability of Clot")
```

##3

```{r,echo=FALSE}
ggplot(sepsis_asc, aes(x=reorder(County,Score.Sepsis),y=Score.Sepsis, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Lowest Sepsis Probability") + xlab("Top Counties") + ylab("Patient Probability of Bacterial Sepsis")
```

##4

```{r,echo=FALSE}
ggplot(sepsis_desc, aes(x=reorder(County,Score.Sepsis),y=Score.Sepsis, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Highest Sepsis Probability") + xlab("Top Counties") + ylab("Patient Probability of Bacterial Sepsis")
```

##5

```{r, echo=FALSE}
ggplot(cath_asc, aes(x=reorder(County,Score.Cath.Infection),y=Score.Cath.Infection, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Lowest Catheder Infection Probability") + xlab("Top Counties") + ylab("Patient Probability of Infection By Catheder")
```

##6

```{r,echo=FALSE}
ggplot(cath_desc, aes(x=reorder(County,Score.Cath.Infection),y=Score.Cath.Infection, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Highest Catheder Infection Probability") + xlab("Top Counties") + ylab("Patient Probability of Infection By Catheder")
```

##7

```{r,echo=FALSE}
ggplot(complications_asc, aes(x=reorder(County,Score.Serious.Complication),y=Score.Serious.Complication, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Lowest Severe Complications in Surgery") + xlab("Top Counties") + ylab("Patient Probability of Complications in Surgery")
```

##8

```{r,echo=FALSE}
ggplot(complications_desc, aes(x=reorder(County,Score.Serious.Complication),y=Score.Serious.Complication, color=factor(State)))+geom_point() + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + ggtitle("Counties With Highest Severe Complications in Surgery") + xlab("Top Counties") + ylab("Patient Probability of Complications in Surgery")
```

## Next Step
- What if we have other data on health?
  *Expanding beyond top 20
